`r params$doc_title`

### 2022-11-14 scripts/MSc.Rmd
several MR methods to calculate theta_X implemented

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

harmonised_dat <- TwoSampleMR::harmonise_data(
  exposure_dat = exposure_dat, 
  outcome_dat = outcome_dat, 
  action = 2  # default
  )

@@

>>>>

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(TwoSampleMR)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
source("MR_lasso.R")  # Slob&Burgess 2020
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

harmonised_dat <- TwoSampleMR::harmonise_data(
  exposure_dat = exposure_dat, 
  outcome_dat = outcome_dat, 
  action = 2  # default
  ) # # used later for analysis with the {TwoSampleMR} package

@@

### Using {TwoSampleMR} package

```{r}
harmonised_dat4TwoSampleMR <- harmonised_dat %>% 
  mutate(id.exposure = "BMI", id.outcome = "GBC")  # https://mrcieu.github.io/TwoSampleMR/articles/perform_mr.html

res <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR)
res <- res %>% mutate(OR = exp(b), CI.lower= exp(b-se), CI.upper=exp(b+se))

res %>% 
  kable(digits = 3, 
        caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI"
        )

#res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
#                                        method_list = c("mr_egger_regression", "mr_ivw"))
```

### Using {MendelianRandomization} package

```{r}
mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
numIV <- nrow(my_data_harm)

if (numIV>2){
        mr.obj = MendelianRandomization::mr_input(
          # exposure
          bx = my_data_harm$beta.bmi, 
          bxse = my_data_harm$se.bmi, 
          # outcome
          by = my_data_harm$beta.gbc, 
          byse = my_data_harm$se.gbc, 
          snps = my_data_harm$SNP,
          exposure = "Body mass index",
          outcome = "Gallbladder Cancer"
        )
        # 1. IVW
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_ivw(mr.obj)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="IVW"] = res$Estimate
        est$se.estimate[est$mr_methods=="IVW"] = res$StdError
        est$time[est$mr_methods=="IVW"] = T1-T0
        rm(res)

        # 2. median
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_median(mr.obj)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="median"] = res$Estimate
        est$se.estimate[est$mr_methods=="median"] = res$StdError
        est$time[est$mr_methods=="median"] = T1-T0
        rm(res)
        # 3. mode
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_mbe(mr.obj)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="mode"] = res$Estimate
        est$se.estimate[est$mr_methods=="mode"] = res$StdError
        est$time[est$mr_methods=="mode"] = T1-T0
        rm(res)
        # 4.MR-PRESSO    ## CAVE: 5 minutes per round, consider only calculating for defined thresholds
        T0 = proc.time()[3]
        res <- MRPRESSO:: mr_presso(BetaOutcome = "beta.gbc", 
                    BetaExposure = "beta.bmi", 
                    SdOutcome = "se.gbc", 
                    SdExposure = "se.bmi", 
                    OUTLIERtest = TRUE, 
                    DISTORTIONtest = TRUE, 
                    seed = 20221114,
                    data = as.data.frame(my_data_harm), ## must be a data.frame
                    NbDistribution = 1000,   # must be changed
                    SignifThreshold = 0.05)
        T1 = proc.time()[3]
        if (!is.na(res$`Main MR results`[2,"Causal Estimate"]) & !is.na(res$`Main MR results`[2,"Sd"])){
                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[2,"Causal Estimate"]
                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[2,"Sd"]
            } else{
                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[1,"Causal Estimate"]
                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[1,"Sd"]
            }
        est$time[est$mr_methods=="PRESSO"] = T1-T0
        rm(res)
        # 5. robust
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_ivw(mr.obj,"random", robust = TRUE)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="Robust"] = res$Estimate
        est$se.estimate[est$mr_methods=="Robust"] = res$StdError
        est$time[est$mr_methods=="Robust"] = T1-T0
        rm(res)
        # 6. MR-Lasso
        T0 = proc.time()[3]
        res = MR_lasso(betaYG = my_data_harm$beta.gbc, betaXG = my_data_harm$beta.bmi, sebetaYG = my_data_harm$se.gbc)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="Lasso"] = res$ThetaEstimate
        est$se.estimate[est$mr_methods=="Lasso"] = res$ThetaSE
        est$time[est$mr_methods=="Lasso"] = T1-T0
        rm(res)
        # 7. Egger
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_egger(mr.obj)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="egger"] = res$Estimate
        est$se.estimate[est$mr_methods=="egger"] = res$StdError.Est
        est$time[est$mr_methods=="egger"] = T1-T0
        rm(res)
        # 8. contamination mixture
        T0 = proc.time()[3]
        res = MendelianRandomization::mr_conmix(mr.obj)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="conmix"] = res$Estimate
        CIlength = res$CIUpper-res$CILower
        if (length(CIlength)>1) print("conmix multimodal")
        est$se.estimate[est$mr_methods=="conmix"] = sum(CIlength)/1.96/2 ## Caution: this may be problematic (Qi&Chatterjee2021)
        est$time[est$mr_methods=="conmix"] = T1-T0
        rm(res)
        # 9. MRMix
              # theta_temp_vec = seq(-0.5,0.5,by=0.01)
        T0 = proc.time()[3]
        res = MRMix::MRMix(betahat_x = my_data_harm$beta.bmi, 
                          betahat_y = my_data_harm$beta.gbc, 
                          sx =my_data_harm$se.bmi, 
                          sy = my_data_harm$se.gbc)
        res_se = MRMix::MRMix_se(betahat_x = my_data_harm$beta.bmi, 
                          betahat_y = my_data_harm$beta.gbc, 
                          sx =my_data_harm$se.bmi, 
                          sy = my_data_harm$se.gbc, 
                          theta = res$theta, # estimate of causal effect, assuming the summary stats are standardized, theta represents increase in mean value of Y in s.d. unit of Y (for cont outcomes) or log-OR of Y (for binary outcomes)
                          pi0 = res$pi0, 
                          sigma2 = res$sigma2
                          )
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="MRMix"] = res$theta
        est$se.estimate[est$mr_methods=="MRMix"] = res_se
        est$time[est$mr_methods=="MRMix"] = T1-T0
        rm(res, res_se)
        # 10. MR-RAPS  
        ## Warning: The estimated overdispersion parameter is very small. Consider using the simple model without overdispersion.
        T0 = proc.time()[3]
        res = mr.raps::mr.raps.overdispersed.robust(b_exp = my_data_harm$beta.bmi, 
                                                    b_out = my_data_harm$beta.gbc, 
                                                    se_exp = my_data_harm$se.bmi, 
                                                    se_out = my_data_harm$se.gbc, 
                                                    loss.function = "huber", 
                                                    k = 1.345, 
                                                    initialization = c("l2"), 
                                                    suppress.warning = FALSE, 
                                                    diagnosis = FALSE, 
                                                    niter = 20, 
                                                    tol = .Machine$double.eps^0.5)
        T1 = proc.time()[3]
        est$estimate[est$mr_methods=="RAPS"] = res$beta.hat
        est$se.estimate[est$mr_methods=="RAPS"] = res$beta.se
        est$time[est$mr_methods=="RAPS"] = T1-T0
        rm(res)
}

est %>% 
  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI")
```
